### 2025-03-18 ESC195 Lecture Notes/main.tex
mar 18

        \intertext{Recall that $\vec{T} = \left | \frac{d\vec{T}}{dt} | \right | \vec{N}$, we have:}

@@

        &= \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\frac{d\phi}{dt}\vec{N} \nonumber \\
        &= \frac{d^2s}{dt^2}\vec{T} + \frac{ds}{dt}\kappa\vec{N}

\appendix
\input{appendix.tex}
\end{document}

>>>>

        \intertext{Recall that $\vec{T} = \left | \frac{d\vec{T}}{dt} \right | \vec{N}$, we have:}

@@

        &= \frac{d^2s}{dt^2}\vec{T} + \frac{d^2s}{dt^2}\kappa\vec{N}
\section{Partial Derivatives}
\subsection{Functions of Several Variables}
\begin{definition}[Functions of Several Variables]
    A function of several variables is a function that maps a point in $\mathbb{R}^n$ to a real number. A function of two variables is given by:
        f(x, y)
\begin{example}[Temperature]
    We have been duelign with funciton that maps a line (like a rod) to a real number (temperature). We can extend this to a function that maps a plane to a real number (temperature of a surface):
    T(x, y)
    Similarly, we can extend this to a function that maps a volume to a real number (temperature of a room):
    T(x, y, z)
\begin{example}[Confinement of Plasma in Magnetic Field in Fusion Reactors]
    Plasma are matter so hot that it is ionized. Hence, they consist in a soup of particles. It is used in fusion reactors. The plasma is confined in a magnetic field. The magnetic force is given by:
        \vec{F} &= q\vec{E}
        \vec{F} &= q\vec{v} \times \vec{B} 
    Let $\vec{B} = B_0\vec{k}$, the trajectory of the particle is given by:
        \vec{F} &= m\vec{r''} \nonumber \\
        \intertext{Let $\omega_L = \frac{qB_0}{m}$ (gyrofrequency), we have:} 
        \vec{v'} = w_L\vec{v} \times \vec{k} \nonumber
        \intertext{Thus, we have the following 3 coupled differential equations:}
        v_x' &= -\omega_Lv_y \nonumber \\
        v_y' &= \omega_Lv_x \nonumber \\
        v_z' &= 0 \nonumber
        \intertext{Differentiating the first equation, we have:}
        v_x'' &= -w_Lv_y' = -w_L^2v_x \nonumber \\
        v_x = A\sin(\omega _Lt) \hspace{0.4in} v_y &= A\cos(\omega _Lt) \hspace{0.4in} v_z = C \nonumber \\
        \vec{v} &= A\sin(\omega_Lt)i + A\cos(\omega_Lt)j + Ck \\ 
        \vec{r} &= \int \vec{v} \, dt = \frac{A}{\omega_L}\cos(\omega_Lt)i - \frac{A}{\omega_L}\sin(\omega_Lt)j + Ct\vec{k}
    Which, we can see is a helix. The current state-of-the art uses a torus to confine the plasma. But this leads to a cicular magnetic field; the gradient of the magnetic field is not constant, this leads to and electric field that can cause the plasma to drift:
\subsection{Partial Derivatives}
\begin{example}[Upper Hemishere]
    Let $f(x, y) = \sqrt{25 - x^2 - y^2}$, we cut the shphere along the surface $y=2$. We can find the slope of the curve at $x=3$. We have:
        f(x, 2) &= \sqrt{25 - x^2 - 4} \\
        f'(x, 2) &= \frac{-x}{\sqrt{25 - x^2 - 4}} \\
        f'(3, 2) &= \frac{-3}{\sqrt{25 - 9 - 4}} = -\frac{3}{\sqrt{12}}
\begin{definition}
    In general, we treat the other variable as a constant. We can find the partial derivative of $f(x, y)$ with respect to $x$ as:
        f_x = \frac{\partial f}{\partial x} = \lim_{h \to 0} \frac{f(x+h, y) - f(x, y)}{h}
    Similarly, we can find the partial derivative of $f(x, y)$ with respect to $y$ as:
        f_y = \frac{\partial f}{\partial y} = \lim_{h \to 0} \frac{f(x, y+h) - f(x, y)}{h}
\paragraph{Notation} We can also write the partial derivative of $f(x, y)$ with respect to $x$ as:
    f_x = \frac{\partial f}{\partial x} = \frac{\partial f}{\partial x} = \frac{\partial}{\partial x}f = D_x f
\paragraph{Note} \textit{We prnounce $\partial$ as `die'}
    Let $f(x, y, z) = e^{xy}\sin(xz)$, we can find the partial derivative of $f(x, y, z)$ with respect to $x$. We have:
        f_x &= \frac{\partial f}{\partial x} = y e^{xy}\sin(xz) + ze^{xy}\cos(xz) \\
        f_y &= \frac{\partial f}{\partial y} = x e^{xy}\sin(xz) \\
        f_z &= \frac{\partial f}{\partial z} = x e^{xy}\cos(xz)
\paragraph{Mixed Partial Derivatives} We can also find the mixed partial derivatives of $f(x, y)$ with respect to $x$ and $y$ as:
    f_{xy} = \frac{\partial^2 f}{\partial y \partial x} = \frac{\partial}{\partial y} \left( \frac{\partial f}{\partial x} \right)
    f_{yx} = \frac{\partial^2 f}{\partial x \partial y} = \frac{\partial}{\partial x} \left( \frac{\partial f}{\partial y} \right)
\begin{theorem}[Clairaut's Theorem]
    If $f(x, y)$ is defined on an open disk $D$ containing the point $(a, b)$ and if the partial derivatives $f_{xy}$ and $f_{yx}$ are continuous on $D$, then $f_{xy}(a, b) = f_{yx}(a, b)$.
\subsection{Directional Derivatives and the Gradient}
\paragraph{Differentiability and Gradients} Using the first principle, we can find the directional derivative of $f(a_1, a_2, \dots, a_n)$ as $\vec{h} \to 0$:
    \lim_{\vec{h} \to 0} \frac{f(\vec{a} + \vec{h}) - f(\vec{a})}{\vec{h}}
\paragraph{A note on $\Pi$ Notation} We can write the product of a series using $\Pi$ notation:
    \prod_{i=1}^n a_i = a_1 \cdot a_2 \cdot \ldots \cdot a_n
\appendix
\input{appendix.tex}
\end{document}

### 2025-03-18 MAT185 Lecture Notes/main.tex
mar 18

\section{Linear combinations}
\begin{definition}[Linear Combination]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$, then:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n
    \end{align*}
    is a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\subsection{Linear Independence, Span, and Basis}
\begin{definition}[Linear independence]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$. If and only if the equation:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \textbf{0}
    \end{align*}
    has \textbf{only} the trivial solution $\alpha_1 = \alpha_2 = \ldots = \alpha_n = 0$.

@@

\section{Determinants}

@@

        \item $f(A) = (\det A )f(I)$
    \end{enumerate}
    is called a determinant function.
\end{definition}
\end{theorem}
\end{document}

>>>>

\section{Linear combinations and Bases}
\begin{definition}[Linear Combination]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$, then:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n
    \end{align*}
    is a linear combination of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\subsection{Linear Independence, Span, and Basis}
\begin{definition}[Linear independence]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$ and $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R}$. If and only if the equation:
    \begin{align*}
        \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n = \textbf{0}
    \end{align*}
    has \textbf{only} the trivial solution $\alpha_1 = \alpha_2 = \ldots = \alpha_n = 0$.

@@

\section{Determinants and Inverses}
\subsection{Determinants}

@@

        \item $f(A) = (\det A )f(I)$
    is called a determinant function.
    If $f$ is alternating on rows and $A$ is a matrix that has two identical rows, then $f(A) = 0$.
\subsection{Constructing Determinants and Inverses}
\begin{definition}[Minor Matrix]
    Let $A \in \mathbb{R}^{n \times n}$, then the minor of $A$ is denoted as $A_{ij}$, which the matrix obtained by removing the $i$-th row and $j$-th column of $A$.
    Let $A = \begin{bmatrix} 1 & 2 & 3 \\ 4 & 5 & 6 \\ 7 & 8 & 9 \end{bmatrix}$, then the minor of $A$ is:
    $$ A_{11} = \begin{bmatrix} 5 & 6 \\ 8 & 9 \end{bmatrix} \quad A_{12} = \begin{bmatrix} 4 & 6 \\ 7 & 9 \end{bmatrix} \quad A_{13} = \begin{bmatrix} 4 & 5 \\ 7 & 8 \end{bmatrix} $$
\paragraph{Find the Determinant from a Row} We can find the determinant of a matrix by expanding along a row. We can expand along any row or column, but it is easier to expand along a row with zeros.
\begin{definition}[Cofactor]
    Let $A \in \mathbb{R}^{n \times n}$, then the cofactor of $A$ is denoted as $C_{ij}$, which is defined as:
        C_{ij} = (-1)^{i+j}\det(A_{ij})
\paragraph{Checkerboard Pattern} We can use the checkerboard pattern to find the determinant of a matrix. We can expand along any row or column, but it is easier to expand along a row with zeros.
    Let $A \in \mathbb{R}^{n \times n}$, then the determinant of $A$ is denoted as $\det(A)$, which is defined as:
        \det(A) = \sum_{j=1}^{n} a_{ij}C_{ij}
    Let $A = \begin{bmatrix} 1 & 2 & 3 \\ 4 & 5 & 6 \\ 7 & 8 & 9 \end{bmatrix}$, then the cofactor of $A$ is:
    $$ C_{11} = \begin{bmatrix} 5 & 6 \\ 8 & 9 \end{bmatrix} \quad C_{12} = \begin{bmatrix} 4 & 6 \\ 7 & 9 \end{bmatrix} \quad C_{13} = \begin{bmatrix} 4 & 5 \\ 7 & 8 \end{bmatrix} $$
\paragraph{Find the Determinant from a Column} We can find the determinant of a matrix by expanding along a column. We can expand along any row or column, but it is easier to expand along a column with zeros.
\paragraph{Using Elementary Row Operations} We can use elementary row operations to simplify the matrix on the row or column we are expanding along (by having zeros).
\begin{theorem}[Determinant w.r.t Elementary Row Operations]
    Consider the following elementary row operations:
        \item \textbf{Swapping two rows}: The determinant of the matrix changes sign due to the alternating property.
        \item \textbf{Multiplying a row by a scalar}: The determinant of the matrix is multiplied by the scalar. This is obvious.
        \item \textbf{Adding a multiple of one row to another row}: The determinant of the matrix remains the same. This is due to the multilinearity property.
\begin{theorem}[Determinant of a Transpose]
    Let $A \in \mathbb{R}^{n \times n}$, then:
        \det(A) = \det(A^T)
\begin{example}[Determinant of a $5\times5$ Matrix]
    Let $A = \begin{bmatrix} 3 & 2 & 1 & 4 & -1 \\ 5 & 4 & 3 & 2 & 1 \\ 2 & 1 & 2 & 3 & 1 \\ -3 & 4 & 1 & 6 & 7 \\ 1 & 2 & 3 & 4 & 5 \end{bmatrix}$. We have:
        \intertext{Step 1: Swap row 1 and row 5}
        \det A &= -\begin{vmatrix} 1 & 2 & 3 & 4 & 5 \\ 5 & 4 & 3 & 2 & 1 \\ 2 & 1 & 2 & 3 & 1 \\ -3 & 4 & 1 & 6 & 7 \\ 3 & 2 & 1 & 4 & -1 \end{vmatrix} \\
        \intertext{Step 2: Get row 2 - 5 start with 0}
        \det A &= -\begin{vmatrix} 1 & 2 & 3 & 4 & 5 \\ 0 & -6 &-12 & -18 & -24 \\ 0 & -3 & -4 & -5 & -9 \\ 0 & 10 & 10 & 18 & 22 \\ 0 & -4 & -8 & -8 & -16 \end{vmatrix} \\
        \intertext{Step 3: Get it to a UTM}
        \det A &= -\begin{vmatrix} 1 & 2 & 3 & 4 & 5 \\ 0 & -6 &-12 & -18 & -24 \\ 0 & 0 & 2 & 4 & 3 \\ 0 & 0 & 0 & 4 & 0 \\ 0 & 0 & 0 & 0 & -3 \end{vmatrix} \\
        \intertext{We can see that each row expansion is just a multiple of the minor below it. Thus, the determinant is the product of the diagonal elements.}
        \det A &= -1 \times -6 \times 2 \times 4 \times -3 = -144
\begin{theorem}[The theorem of Determinants and Inverse]
    Let $A \in \mathbb{R}^{n \times n}$, then the following properties hold:
        \item $A$ is invertible.
        \item $\det(A) \neq 0$.
        \item $\text{RREF}(A) = I$.
        \item $A$ is a product of elementary matrices.
        \item $\det(A^{-1}) = \frac{1}{\det(A)}$.
        \item $\text{rank}(A) = n$.
\section{Eigen and Diagonalization}
\subsection{Eigenvalues and Eigenvectors}
\begin{definition}[Eigenvalues and Eigenvectors]
    Let $A \in {^n\mathbb{R}^n}$, then $\lambda$ is an eigenvalue of $A$ if there exists a \textbf{non-zero} vector $\textbf{v}$ such that:
        A\textbf{v} = \lambda \textbf{v}
    The vector $\textbf{v}$ is called an eigenvector of $A$.
\begin{theorem}[Properties of Eigenvalues and Eigenvectors]
    Let $A \in {^n\mathbb{R}^n}$, then the following properties hold:
        \item The eigenvalues of $A$ are the roots of the characteristic equation:
            \det(A - \lambda I) = 0
        \item The eigenvectors of $A$ are the solutions to the system of equations:
            (A - \lambda I)\textbf{v} = \textbf{0}
\paragraph{Note} If $A\textbf{v} = \textbf{0}$, then $\textbf{v}$ is an eigenvector of $A$ with eigenvalue $\lambda = 0$.
\begin{example}[Eigenvalues and Eigenvectors]
    Let $A = \begin{bmatrix} 1 & 2 \\ 2 & 1 \end{bmatrix}$, then the characteristic equation is:
        \det(A - \lambda I) &= \begin{vmatrix} 1-\lambda & 2 \\ 2 & 1-\lambda \end{vmatrix} \\
        &= (1-\lambda)(1-\lambda) - 4 \\
        &= \lambda^2 - 2\lambda - 3 = 0
    The roots of the characteristic equation are $\lambda = 3, -1$. The eigenvectors are:
        \text{For } \lambda = 3: \begin{bmatrix} -2 & 2 \\ 2 & -2 \end{bmatrix}\textbf{v} &= \textbf{0} \implies \textbf{v} = \begin{bmatrix} 1 \\ 1 \end{bmatrix} \\
        \text{For } \lambda = -1: \begin{bmatrix} 2 & 2 \\ 2 & 2 \end{bmatrix}\textbf{v} &= \textbf{0} \implies \textbf{v} = \begin{bmatrix} 1 \\ -1 \end{bmatrix}
    Let $A \in {^n\mathbb{R}^n}$ and has eigen value 2 with multiplicity 2, and eigen value 3 with multiplicity 1. Let $x, y, z$ be eigenvectors of $A$ corresponding to the eigenvalues 2, 2, 3 respectively. Then:
        \item \textbf{2x is an eigenvector of A}: Since $A(2x) = 2Ax = 2(2x)$, then $2x$ is an eigenvector of $A$ with eigenvalue 2.
        \item \textbf{x + y is an eigenvector of A}: Since $A(x + y) = Ax + Ay = 2x + 2y = 2(x + y)$, then $x + y$ is an eigenvector of $A$ with eigenvalue 2.
        \item \textbf{x + z is an not an eigenvector of A}: Since $A(x + z) = Ax + Az = 2x + 3z \neq \lambda(x + z)$, then $x + z$ is not an eigenvector of $A$.
\subsection{Eigenspace and Multiplicity}
\begin{definition}[Eigenspace]
    Let $A \in {^n\mathbb{R}^n}$, then the eigenspace of $A$ corresponding to the eigenvalue $\lambda$ is denoted as $E_\lambda$, which is defined as:
        E_\lambda = \{ \textbf{v} \in \mathbb{R}^n \mid A\textbf{v} = \lambda \textbf{v} \}
    , which is the union of all eigenvectors of $A$ corresponding to the eigenvalue $\lambda$ and ${\textbf{0}}$.
\begin{theorem}[Eigenspace is a Subspace]
    Let $A \in {^n\mathbb{R}^n}$, then the eigenspace of $A$ corresponding to the eigenvalue $\lambda$ is a subspace of $\mathbb{R}^n$ with dimension equal to the multiplicity of $\lambda$ ($\ge 1$).
    This is obvious as eigenspace is a span of eigenvectors, and the span of eigenvectors is a subspace. To proof it, we use the subspace test:
        \item \textbf{Non-emptiness}: Since $\textbf{0}$ satisfies $A\textbf{0} = \lambda \textbf{0}$, then $\textbf{0} \in E_\lambda$.
        \item \textbf{Closure under addition}: Let $\textbf{v}_1, \textbf{v}_2 \in E_\lambda$, then $A(\textbf{v}_1 + \textbf{v}_2) = A\textbf{v}_1 + A\textbf{v}_2 = \lambda \textbf{v}_1 + \lambda \textbf{v}_2 = \lambda(\textbf{v}_1 + \textbf{v}_2)$, then $\textbf{v}_1 + \textbf{v}_2 \in E_\lambda$.
        \item \textbf{Closure under scalar multiplication}: Let $\textbf{v} \in E_\lambda$, then $A(c\textbf{v}) = cA\textbf{v} = c\lambda \textbf{v} = \lambda(c\textbf{v})$, then $c\textbf{v} \in E_\lambda$.
\begin{theorem}[Intersection of Eigenspaces]
    Let $A \in {^n\mathbb{R}^n}$, then the intersection of eigenspaces of $A$ corresponding to distinct eigenvalues is $\{\textbf{0}\}$, that is, for eigenvalues $\lambda_1, \lambda_2$, if $\lambda_1 \neq \lambda_2$, then $E_{\lambda_1} \cap E_{\lambda_2} = \{\textbf{0}\}$.
\begin{definition}[Algebraic Multiplicity (AM)]
    Let $A \in {^n\mathbb{R}^n}$, then the algebraic multiplicity of an eigenvalue $\lambda$ is the number of times $\lambda$ appears as a root of the characteristic equation.
\begin{definition}[Geometric Multiplicity (GM)]
    Let $A \in {^n\mathbb{R}^n}$, then the geometric multiplicity of an eigenvalue $\lambda$ is the dimension of the eigenspace of $\lambda$.
\begin{theorem}[Multiplicity of Eigenvalues]
    Let $A \in {^n\mathbb{R}^n}$, then the algebraic multiplicity of an eigenvalue is greater than or equal to the geometric multiplicity of the eigenvalue, that is:
        \text{AM}(\lambda) \ge \text{GM}(\lambda) \ge 1
\begin{theorem}[Product of Eigenvalues]
    Let $A \in \mathbb{R}^{n \times n}$, then the product of the eigenvalues of $A$ is equal to the determinant of $A$:
        \prod_{i=1}^{n} \lambda_i = \det(A)
\end{document}

### 2025-03-18 MSE160 Lecture Notes/main.tex
mar 18

>>>>

\subsection{Matter-Energy Interactions}
\paragraph{Photo Wavelength-Frequency} The wavelength of a photon is inversely proportional to its frequency. The relationship is given by:
    c = f\lambda
\paragraph{Energy of a Photon} The energy of a photon is given by:
    E = hf = \frac{hc}{\lambda}
\paragraph{Visible Light} Visible light is the range of wavelengths that the human eye can see. It is between 400 nm and 700 nm (3eV to 2eV).
\paragraph{Bohr's Model} Bohr's model is a model of the atom that describes the electron as a particle in a circular orbit around the nucleus. It discribes electrons in well-defined qunatized orbits. Thus, electrion promotes and demotes between orbitals, creating discrete spectra.
\paragraph{Lmitation of Bohr's Model} Bohrs model cannot captures the energy levels and orbitals of atoms with more than one electron. It also cannot explain the fine structure of the spectra.
\paragraph{Wave-Mechanical Model} The wave-mechanical model describes the electron as a wave. It describes the electron as a wavefunction, $\psi$, that describes the probability of finding the electron at a certain position. 
\paragraph{Quantum Numbers} The wave-mechanical model uses quantum numbers to describe the electron. The quantum numbers are:   
    \item \textbf{Principal Quantum Number} $n = 1,2,3,\dots$ - Describes the energy level of the electron (size of the orbital). \\
    (note) $n$ is sometimes given by $n = K, L, M, \dots$
    \item \textbf{Angular Momentum Quantum Number} $l = 0,1,2,\dots,n-1$ - Describes the shape of the orbital.
    (note) Those are the $s,p,d,f$ orbitals.
        \item \textbf{$l=0$ or $s$} Spherical
        \item \textbf{$l=1$ or $p$} Dumbbell
        \item \textbf{$l=2$ or $d$} Four-leaf clover
        \item \textbf{$l=3$ or $f$} Complex
    \item \textbf{Magnetic Quantum Number} $m_l = -l, -l+1, \dots, 0, \dots, l-1, l$ - Describes the orientation of the orbital.
    \item \textbf{Spin Quantum Number} $m_s = \pm \frac{1}{2}$ - Describes the spin of the electron.
\paragraph{Relative Energy Levels} The energy levels of the orbitals are given by:
    E = -\frac{Z^2}{n^2} \text{Ry}
\paragraph{Electrion Configuration} The electron configuration is the distribution of electrons in the orbitals. It is given by:
    \text{Element} = [n]l^m
(Example) The electron configuration of Carbon is $1s^2 2s^2 2p^2$. That for iron is $1s^2 2s^2 2p^6 3s^2 3p^6 4s^2 3d^6$.
\paragraph{Stable Octet} The $ns^2 np^6$ is a stable configuration that forms the basis for the primary bonds. Atoms tend to gain, lose or share electrons to achieve this configuration. As we know, if we share electrons, we form covalent bonds; if we lose or gain electrons, we form ionic bonds.
\paragraph{Half filled and Fully Filled Orbitals} Half filled and fully filled orbitals are more stable than partially filled orbitals. Thus at $d^5$ and $d^{10}$, the $d$ orbitals are more stable and it tends to lend electron from the $s$ orbital.\\
(Example) Chromium is $[Ar] 3d^5 4s^1$ and Copper is $[Ar] 3d^{10} 4s^1$.
\paragraph{Pauli Exclusion Principle} The Pauli Exclusion Principle states that no two electrons can have the same set of quantum numbers. 
\paragraph{Tightly Packed Atoms} When atoms are packed, due to the pauli exclusion principle, it creates a bend of energy levels of the same orbital. In ground state, the electrons will fill the lower energy levels first.
\subsection{Band Theory}
\paragraph{Conductivity} At room temperature, the thermal promotion within the bend makse it possible for conduction in conductors.
\paragraph{Overlaping Energy Level} In some atoms, the bend of energy level gets larger and overlap with the energy level of the next orbital. 
\paragraph{Conductive Band and Valence Band} The conductive band is the band of energy levels that are above the valence band. The valence band is the band of energy levels that are filled with electrons. There is a band gap between the two bands, the size of the band gap determines the conductivity of the material.
\paragraph{Conductors} Conductors have a no band gap. The valence band and conductive band overlap. Thus, electrons can easily move between the two bands.
\paragraph{Insulators} Insulators have a large band gap. The valence band and conductive band are far apart. Thus, electrons cannot easily move between the two bands.
\paragraph{Semiconductors} Semiconductors have a small/moderate band gap. The valence band and conductive band are close together. Thus, electrons can move between the two bands with the addition of energy.
\textbf{For This Coruse, we take the band gap for semiconductors to be $\le 4$ eV.}
\paragraph{Optically transparency} A material would be optically transparent if the band gap is larger than the energy of the visible light (3 eV).
\subsection{Silicon}
\paragraph{Intrinsic Silicon} Intrinsic silicon is pure silicon. When electron is promoted to the conduction band, it leaves a hole in the valence band. The hole can move through the lattice as a postive charge. The electron can move through the lattice as a negative charge. The electron-hole pair is called an exciton. For intrinsic silicon, the number of holes is equal to the number of electrons. Also, the charge carrier contcentration is given by:
    \frac{n=p}{a^3}
For silicon, at room temperature, the charge carrier concentration is $10^{16}$ m$^{-3}$.
\paragraph{Doping} Doping is the addition of impurities to a semiconductor. They are point defects in the lattice. Doping is used to increase the conductivity of the semiconductor. There are two types of doping:
    \item \textbf{N-Type Doping} N-type doping is the addition of impurities that increase the number of electrons. The extra electron is in the higher end of the band gap. The impurities are called donor impurities. The donor impurities are typically group 15 elements. The donor impurities are negatively charged.
    \item \textbf{P-Type Doping} P-type doping is the addition of impurities that increase the number of holes. The impurities are called acceptor impurities. The acceptor impurities are typically group 13 elements. The acceptor impurities are positively charge the the hole is on the higher end of the valence band and to be promoted to the lower end of the band gap.
